CONEM2024-0025\\ ABC MULTICRITERIA CLASSIFICATION WITH ANALYTICAL HIERARCHY PROCESS: HOW TO FORCE CONSISTENCY OF PAIRWISE COMPARISONS MATRIX

### 2024-02-16 accepted/CONEM2024/FCPCOM/CONEM2024.tex
work in progress - writing introduction - not finished yet

	The importance of multicriteria classification for inventory management has been widely emphasized in scientific literature and several multicriteria methods have been used, being the ABC multicriteria classification with Analytical Hierarchy Process one of the most common techniques applied for this purpose \citep{Flores1992, AltayGuvenir1998, Lolli2014, Balaji2014}. Few recent works talk about multicriteria classification applied to strategic planning \citep[\emph{e.g.},][]{BarbosaDePaula2022, Pereira2023, MarianoRibeiro2023}.

	\citet{Odu2019} emphasize that, when aplying some multicriteria method, it is important to pay particular attention to the objectivity factors of criteria weights. The author provides an important overview of different weighting methods applicable to multicriteria techniques. In the present work, we opted for the multicriteria ABC classification and adopted the Analytic Hierarchy Process \citep{Saaty1987} to assign the weights to the different criteria. It is important to note that the Analytic Hierarchy Process derives the weights of the criteria based on the eigenvector of a pairwise comparison matrix which is constructed manually according to an empirical evaluation. According to \cite{Saaty1987}, this matrix must be antisymmetric and must have a consistency rate of less than 0.1.  The autor reports that a software package supporting the AHP, called Expert Choice, for the IBM PC, was used in his work for making eigenvector calculations and guide the decision maker to improve matriz inconsistency if needed. During the execution of the work we present in this paper, we found that there is a strong difficulty in for calculation of the eigenvector, as well as for the construction of consistent pairwise comparison matrices and we had no access to the Expert Choice software package, so we developed an open source COPSolver library for Analitic Hierarch Process and multicriteria ABC classification based on a new a method for forcing the consistency of such matrices. We present this method in this paper. The Eigen library from eigen.tuxfamily.org was used for eigenvector calculation. Also we identified that the assignment of weights to ABC classification can be done using different criteria and we observed that such assignment generates a bias in the multicriteria ABC classification. Therefore we also developed a new methodology for assigning weights in a balanced way for ABC classification and we proposed a new multicriteria ABC classification method, that we aplyed on COPSolver library. We also presente both methods on this paper and an in-depth analysis of the results found by the library we developed for COPSolver.

@@

    Inventory classification problem is also a decision problem since we want to decide what is the better system and criteria for clustering the different inventory itens.

In our study, we observed that one of the crucial points for correctly applying multicriteria ABC classification is correctly attributing the weights for the initial ABC classification.

>>>>

	\citet{Odu2019} emphasize that, when aplying some multicriteria method, it is important to pay particular attention to the objectivity factors of criteria weights. The author provides an important overview of different weighting methods applicable to multicriteria techniques. The importance of multicriteria classification for inventory management has been widely emphasized in scientific literature and several multicriteria methods have been used, being the ABC multicriteria classification with Analytical Hierarchy Process one of the most common techniques applied for this purpose \citep{Flores1992, AltayGuvenir1998, Lolli2014, Balaji2014}. Few recent works talk about multicriteria classification applied to strategic planning \citep[\emph{e.g.},][]{BarbosaDePaula2022, Pereira2023, MarianoRibeiro2023}.

	In the present work, we opted for the multicriteria ABC classification and adopted the Analytic Hierarchy Process (AHP) \citep{Saaty1987} to assign the weights to the different criteria. It is important to note that the Analytic Hierarchy Process derives the weights of the criteria based on the eigenvector of a pairwise comparison matrix which is constructed manually according to an empirical evaluation. According to \cite{Saaty1987}, this matrix must be antisymmetric and must have a consistency rate of less than 0.1.  The autor reports that a software package supporting the AHP, called Expert Choice, for the IBM PC, was used in his work for making eigenvector calculations and guide the decision maker to improve matriz inconsistency if needed. During the execution of the work we present in this paper, we found that there is a strong difficulty in calculation of the eigenvector, as well as for the construction of consistent pairwise comparison matrices and we had no access to the Expert Choice software package, so we developed an open source COPSolver library for Analitic Hierarch Process and multicriteria ABC classification based on a simple algorithm for forcing the consistency of such matrices. We present this algorithm in this paper. The Eigen library from eigen.tuxfamily.org was coupled to COPSolver for eigenvectors calculation of the matrices. Also we identified that the assignment of weights to ABC classification can be done using different criteria and we observed that such assignment generates a bias in the multicriteria ABC classification. Therefore we also developed a new methodology for assigning weights in a balanced way for ABC classification and we proposed a new multicriteria weighting equation for the multicriteria classification COPSolver library. We also present both propositions on this paper and an in-depth analysis of the results found by the new library we developed for COPSolver.

@@

As discussed earlier, in our study, we observed that one of the crucial points for correctly applying multicriteria ABC classification is correctly attributing the weights for the initial ABC classification.
